Supplementary Information

### 2021-06-01 writing/appendix_a.Rmd
Paper v2.1

---
title: Supplementary Information
short_title: COVID-19 Prescribing
bibliography      : ["r-references.bib", "../references/library.bib"]
ouput: html_document
---

@@

```{r, analysis-preferences}
# Seed for random number generation
set.seed(32)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed, cache = TRUE)
kable_opts <- options(knitr.kable.NA = "-")
```

@@

* We did not assume that the existing processes would be the same for all drugs or all GP practices, therefore, we investigated the fit of several different time series models to the pre-COVID data. 
* We used the `fable` [@R-fable], `fasster` [@R-fasster], and the `fable.prophet` [@R-fable.prophet] packages to conduct the time series analyses.
  * Simple methods
      * Naïve (with and without drift)
        * $\hat{y}_{T+h|T} = y_T$
      * Seasonal naïve (with and without drift)
        * $\hat{y}_{T+h|T} = y_{T-m(k+1)}$ where m = seasonal period and k is the complete years in $T+h$
  * STL Decomposition model [@Cleveland1990]
    * Fixed and variable seasonality
  * Time series linear model
    * $y_t = \beta_0 + \beta_1x_t + \varepsilon_t$
  * Autoregressive integrated moving average [ARIMA; @Box2015]
    * Using a version of the Hyndman-Khandakar algorithm [@Hyndman2008]
  * Exponential smoothing [@Holt2004; @Winters1960]
    * Simple exponential smoothing
    * Holt-Winters Additive Model [@Chatfield1978]
  * Forecasting with Additive Switching of Seasonality, Trends and Exogenous Regressors [fasster; @R-fasster]
    * State space model
  * Prophet [@Taylor2018]
  * Combination models [cf. @Bates1969; @Clemen1989; @Thomson2019]
* Prescribing quantities log transformed and forecasts use median values to reduce bias that back transformation would introduce when using the mean
* Using Jan 2015 to Dec 2019 data, we fitted several different models and assessed their accuracy using a cross-validated process to reduce the likelihood of overfitting models to the data.
  * Started with 36 months of data, used 6-month horizon as this was the horizon we would be using for the forecasts, 3-month step (to keep run-times manageable)
    * Only interested in one forecast horizon: six-months

## All prescribing

```{r, all-prescribing-model-accuracy-hist, fig.cap="All prescribing, model accuracy", eval=FALSE}
p1 <- all_prescribing_fable_fits %>% 
  na.omit() %>% 
  ggplot(aes(x = reorder(.model, RMSE, sort), y = RMSE)) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 90))

p2 <- all_prescribing_fable_fits %>% 
  na.omit() %>% 
  ggplot(aes(x = reorder(.model, winkler, sort), y = winkler)) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 90))

cowplot::plot_grid(p1, p2)
```

```{r, all-prescribing-cv-plot, fig.cap="All prescribing, best model CV fit", eval=FALSE}
all_prescribing_best_cv_model <- 
  augment(all_prescribing_fable_models) %>% 
  as_tibble() %>% 
  filter(.model == all_prescribing_best_model$.model)

all_prescribing_best_cv_model %>% 
  ggplot(aes(x = datename)) +
  geom_line(aes(y = n/1e+06, colour = "Data")) +
  geom_line(aes(y = .fitted/1e+06, colour = "Fitted", lty = as.factor(.id))) +
  labs(lty = "CV iteration")
```

```{r, all-prescribing-forecast-plot, warning=FALSE, fig.cap="All prescribing, best model fit and forecast.", eval=FALSE}
plan(cluster)
all_prescribing %>% 
  ggplot(aes(x = datename)) +
  geom_line(aes(y = n, col = "Observed", fill = "Observed")) +
  geom_line(aes(y = med, col = "Forecast", fill = "Forecast"), 
            data = all_prescribing_covid_forecast) +
  geom_ribbon(aes(ymin = lb_80, ymax = ub_80, 
                  fill = "Forecast"), alpha = .2,
              data = all_prescribing_covid_forecast) +
  geom_ribbon(aes(ymin = lb_95, ymax = ub_95, 
                  fill = "Forecast"), alpha = .3,
              data = all_prescribing_covid_forecast) +
  # geom_line(aes(y = fitted, col = "Fitted", fill = "Fitted"), 
  #           data = fableModels(all_prescribing_pre_cv19) %>% 
  #             augment() %>% 
  #             as_tibble() %>% 
  #             filter(.model == all_prescribing_best_model$.model) %>% 
  #             group_by(datename) %>% 
  #             summarise(fitted = median(.fitted, na.rm = TRUE))) +
  geom_vline(xintercept = ymd("2020-03-01"), lty = 3) +
  coord_cartesian(xlim = c(dmy("01-11-2018"), dmy("01-08-2020"))) +
  scale_x_yearmonth(breaks = "6 months", minor_breaks = "1 months")+
  xlab("Date") +
  ylab("Mean daily prescribing volume")

plan(sequential)
```

```{r, all-prescribing-pct-change-plot, fig.cap="Percentage difference in all prescribing.", eval=FALSE}
all_prescribing %>% 
  inner_join(all_prescribing_covid_forecast, by = "datename") %>% 
  ggplot(aes(x = datename, y = (100*n/med)-100)) +
  geom_line() +
  geom_hline(yintercept = 0, lty = 3)
```

@@

n_practices %>% 
  left_join(missing_raw_data, by = "drug") %>% 
  left_join(poor_forcasts, by = "drug") %>% 
  rowwise() %>% 
  mutate(mape_over_50 = if_else(is.na(mape_over_50), as.integer(0), mape_over_50),
         total_removed = miss_either + mape_over_50) %>% 
  knitr::kable(booktabs = TRUE, linesep = "",
               caption = "Summary of GP practices removed for each drug.") %>% 
  footnote(general = "The total is not always equal to the row sum, as a single practice may be missing both data in covid months and more than 10\\% of data in the non-covid months, but will only be excluded once.",
           threeparttable = TRUE)

n_practices %>% 
  summarise(sum(n))

length(unique(gp$practice_id))
```

@@

```{r, cv-model-accuracy-by-drug-tab}
cv_model_files %>% 
  map_dfr(.id = "drug", .f = read_rds) %>% 
  rankModels() %>% 
  group_by(drug, regional_unit) %>% 
  slice(1, .preserve = TRUE) %>% 
  filter(MAPE < 50) %>% 
  group_by(drug) %>% 
  summarise(n_practice_retained = n(),
            mape_mean = mean(MAPE),
            mape_sd = sd(MAPE)) %>% 
  kable(caption = "Accuaracy of cross-validated models retained for forecasting.",
               digits = 2)
```

@@

```{r, oac-gen-trend, eval=FALSE}
oac <- read_rds(data_files[9])

oac %>% 
  mutate(warfarin = if_else(bnfchem == "0208020V0", TRUE, FALSE)) %>% 
  group_by(warfarin, datename) %>% 
  summarise(quantity = sum(quantity), .groups = "drop") %>% 
  pivot_wider(id_cols = datename, values_from = quantity, names_from = warfarin) %>% 
  mutate(prop = `TRUE` / (`TRUE` + `FALSE`),
         datename = yearmonth(dmy(paste0("28-", datename)))) %>% 
  tsibble() %>% 
  autoplot(prop) +
  xlab("Date") +
  ylab("Warfarin as a proportion of all oral anticoagulant prescribing")

>>>>

---
title: "Supplementary Information"
short_title: COVID-19 Prescribing
bibliography:
- r-references.bib
- ../references/library.bib
output: pdf_document
ouput: pdf_document
---

@@

# Seed for random number generation
set.seed(32)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed, cache = TRUE, echo = FALSE)
kable_opts <- options(knitr.kable.NA = "-")
```

@@

* We did not assume that the existing processes would be the same for all drugs or all GP practices, therefore, we investigated the fit of several different time series models to the pre-COVID data. 
* We used the `fable` [@R-fable] and `fasster` [@R-fasster] packages to conduct the time series analyses.
  * Simple methods
      * Naïve (with and without drift)
        * $\hat{y}_{T+h|T} = y_T$
      * Seasonal naïve (with and without drift)
        * $\hat{y}_{T+h|T} = y_{T-m(k+1)}$ where $m$ = seasonal period and $k$ is the complete years in $T+h$
  * STL Decomposition model [@Cleveland1990]
    * Fixed and variable seasonality
  * Time series linear model
    * $y_t = \beta_0 + \beta_1x_t + \varepsilon_t$
  * Autoregressive integrated moving average [ARIMA; @Box2015]
    * Using a version of the Hyndman-Khandakar algorithm [@Hyndman2008]
  * Exponential smoothing [@Holt2004; @Winters1960]
    * Simple exponential smoothing
    * Holt-Winters Additive Model [@Chatfield1978]
  * Forecasting with Additive Switching of Seasonality, Trends and Exogenous Regressors [fasster; @R-fasster]
    * State space model
  * Combination models [cf. @Clemen1989; @Thomson2019]
* Prescribing quantities were log transformed and the forecasts use median estimates of prescribing to reduce bias that back transformation would introduce when using the mean
* Using January 2015 to December 2019 data, we fitted several different models and assessed their accuracy using a cross-validated process to reduce the likelihood of overfitting models to the data.
  * Started with 36 months of data, used 6-month horizon as this was the horizon we would be using for the forecasts, 3-month step (to keep run-times manageable)
    * Only interested in one forecast horizon: six-months

@@

n_practices %>% 
  left_join(missing_raw_data, by = "drug") %>% 
  left_join(poor_forcasts, by = "drug") %>% 
  rowwise() %>% 
  mutate(mape_over_50 = if_else(is.na(mape_over_50), as.integer(0), mape_over_50),
         total_removed = miss_either + mape_over_50) %>% 
  knitr::kable(booktabs = TRUE, linesep = "",
               caption = "Summary of GP practices removed for each drug.") %>% 
  footnote(general = "The total is not always equal to the row sum, as a single practice may be missing both data in covid months and more than 10\\% of data in the non-covid months, but will only be excluded once.",
           threeparttable = TRUE)
```

@@

```{r, cv-model-accuracy-by-drug-tab, eval=FALSE}
cv_model_files %>% 
  map_dfr(.id = "drug", .f = read_rds) %>% 
  rankModels() %>% 
  group_by(drug, regional_unit) %>% 
  slice(1, .preserve = TRUE) %>% 
  filter(MAPE < 50) %>% 
  group_by(drug) %>% 
  summarise(n_practice_retained = n(),
            mape_mean = mean(MAPE),
            mape_sd = sd(MAPE)) %>% 
  kable(caption = "Accuaracy of cross-validated models retained for forecasting.",
               digits = 2)
```

@@

```{r, oac-gen-trend, fig.cap="Warfarin as a percentage of all oral anticoagulant prescribing in Wales. Pink shading covers duration of Wales' \"stay at home\" order."}
oac <- read_rds(data_files[12])

oac %>% 
  mutate(warfarin = if_else(bnfchem == "0208020V0", TRUE, FALSE)) %>% 
  group_by(warfarin, datename) %>% 
  summarise(quantity = sum(quantity), .groups = "drop") %>% 
  pivot_wider(id_cols = datename, values_from = quantity, names_from = warfarin) %>% 
  mutate(prop = `TRUE` / (`TRUE` + `FALSE`),
         datename = yearmonth(dmy(paste0("28-", datename)))) %>% 
  tsibble(index = datename) %>% 
  ggplot(aes(x = datename)) + 
  geom_line(aes(y = prop))+
  geom_vline(aes(xintercept = dmy("01-03-2020")), lty = 2, col = "red") +
  geom_rect(aes(xmin = dmy("01-03-2020"), xmax = dmy("31-07-2020"), 
                ymin = -Inf, ymax = Inf),
            fill = 'pink', col = 0,alpha = 0.03) +
  xlab("Date") +
  ylab("Warfarin as a proportion of all oral anticoagulant prescribing")
